\section{Functionality of Zeek}
Zeek stands out as a versatile platform for network security and analysis for several compelling reasons:
\begin{enumerate}
    \item \textbf{Focused on Network Data:} Specialization in collecting and analyzing network data, a critical component for security teams.
    
    \item \textbf{Diverse Data Collection:} Support for various data types including transaction data, extracted content, and alert data, as defined by the network security monitoring paradigm.
    
    \item \textbf{Rich Transaction Logs:} Default generation of detailed, high-fidelity, and richly-annotated transaction logs, offering a neutral view of network activities.
    
    \item \textbf{File Extraction Capabilities:} Robust features for carving files from network traffic, facilitating further analysis with various tools.
    
    \item \textbf{Alert Mechanism:} Provision of an alert system through the notice mechanism, enabling judgment on network traffic.
    
    \item \textbf{Customization and Extensibility:} Availability of a domain-specific scripting language for tailoring functionalities to specific analysis needs.
    
    \item \textbf{Cost-Effective Deployment:} Ability to run on commodity hardware, making it an accessible solution for organizations of all sizes.
    
    \item \textbf{Comprehensive Analysis and Detection:} Inclusion of built-in functionalities for a wide array of analysis and detection tasks, enhancing its utility as an analysis tool.
\end{enumerate}
