\subsection{Optimal Locations for Zeek Deployment}
For effective network monitoring and security analysis, it is crucial to deploy Zeek in strategic locations within the network infrastructure. Below are key points outlining the recommended deployment locations for Zeek:

\begin{enumerate}
    \item \textbf{At Network Perimeters:} Deploy Zeek at the boundaries of the network to monitor all incoming and outgoing traffic, enabling the identification of external threats and unauthorized data exfiltration.
    
    \item \textbf{Core Network Infrastructure:} Position Zeek near central routing and switching devices to capture a broad spectrum of internal network traffic, providing insights into the core network's interactions.
    
    \item \textbf{Data Center Entrances:} Install Zeek at the entry and exit points of data centers to oversee traffic to critical servers and storage systems, safeguarding sensitive information and key assets.
    
    \item \textbf{Network Segmentation Gateways:} Place Zeek at the junctions between different network segments or VLANs to monitor cross-segment traffic, essential for detecting lateral movements and internal threats.
    
    \item \textbf{VPN Concentrators and Access Points:} Deploy Zeek where VPN connections converge or near wireless access points to analyze encrypted and wireless traffic, respectively, ensuring secure remote access and wireless communication.
    
    \item \textbf{Demilitarized Zones (DMZs):} Position Zeek within or adjacent to DMZs to scrutinize traffic between the internet and internal network, protecting services exposed to the outside world.
    
    \item \textbf{Branch Offices:} Extend Zeek deployment to branch or remote offices to monitor localized traffic, ensuring comprehensive network security coverage across all organizational units.
    
    \item \textbf{High Availability Clusters:} In critical network environments, deploy Zeek in a high availability configuration with load balancing to ensure continuous monitoring and redundancy, preventing single points of failure.
\end{enumerate}
