\section{Introduction}
\subsection{Overview}
Zeek is a passive, open-source network traffic analyzer extensively utilized in the realm of network security monitoring (NSM) and a plethora of other traffic analysis tasks. Originally developed for security investigations, Zeek's capabilities transcend to performance measurement and network troubleshooting, offering users a comprehensive view of their network activity through detailed logs and analysis features.
\subsubsection{Comprehensive Logging}
One of the primary advantages of using Zeek is its ability to produce extensive logs that capture a detailed record of network activities. These logs include:
\begin{itemize}
    \item Application-layer transcripts such as HTTP sessions, with details like URIs, headers, MIME types, and server responses.
    \item DNS requests and responses.
    \item SSL certificate information.
    \item Key content from SMTP sessions.
\end{itemize}
Zeek formats these logs in structured, tab-separated, or JSON files, making them ideal for post-processing with external software or integration with databases and SIEM products for further analysis.


\subsubsection{Built-in Analysis and Detection Capabilities}
Zeek is not just about logging; it offers built-in functions for various analysis and detection tasks, such as:
\begin{itemize}
    \item Extracting files from HTTP sessions.
    \item Detecting malware by interfacing with external registries.
    \item Identifying vulnerable software versions and popular web applications on the network.
    \item Detecting SSH brute-force attacks and validating SSL certificate chains.
\end{itemize}
These functionalities are readily available "out of the box," enhancing Zeek's utility as a network analysis tool.

\subsubsection{Customization and Extensibility}
Zeek stands out for its customizability and extensibility, thanks to its domain-specific scripting language. This language allows users to:
\begin{itemize}
    \item Express arbitrary analysis tasks.
    \item Write custom code for unique analysis requirements, extending the pre-built "standard library" that comes with Zeek.
\end{itemize}

This feature positions Zeek as a versatile tool that can adapt to various network monitoring and analysis needs beyond its default offerings.

\subsubsection{High-Performance and Scalability}
Designed to run on commodity hardware, Zeek presents a cost-effective alternative to proprietary network monitoring solutions. Its performance and scalability make it suitable for:
\begin{itemize}
    \item High-speed, high-volume network monitoring, including 10GE and 100GE networks.
    \item Deployment in large-scale environments using Zeek Clusters for load-balancing and centralized management.
\end{itemize}
 Zeek's scalability is one of its core strengths, allowing it to accommodate growing network demands without compromising on performance.
 \subsubsection{Comparison with Other Tools}
 While Zeek excels in generating high-fidelity network logs and analysis, it is distinct from:

 \begin{itemize}
     \item Signature-based IDS systems like Suricata, as Zeek adopts a broader approach to detecting malicious activity through scripting.
     \item Protocol analyzers like Wireshark, which focus on detailed frame-level traffic representation, unlike Zeek's focus on compact, high-fidelity logs.
     \newline
 \end{itemize}
Zeek offers a balanced and efficient solution for network traffic analysis, combining comprehensive logging, built-in analysis capabilities, and extensive customization options. Its adaptability to both small-scale and large-scale environments, along with its open-source nature, makes Zeek a valuable tool for a wide range of users looking to gain deeper insights into their network activity.